---
title: Commuting Mode Choice based on distance to the next train station
author: Marijana Maric and Pablo Bischofberger
output: html
bibliography: pted_bib.bib
csl: harvard-educational-review.csl
format:
  html:
    code-fold: true
    toc: true
    toc-depth: 4
    toc-location: left
    toc-expand: true
    
warning: false
message: false
---

<!-- You can add  your R Code with Code chunks-->

```{r}
#| echo: false
#| warning: false
#| message: false

# You can set chunk options individually per code chunk, as was done with this
# code chunk.

# echo: false           hides the code from the generated output
# warning: false        hides warnings from the generated output
# message: false        hides messages from the generated output

# To apply the setting for all code chunks, add the options to the yaml header of the document (in between the ---) without the preceeding "#|".

```

## Abstract

This study examines the relationship between the distance from home to the train station and the transportation mode used. The analysis of trajectory data seeks to identify patterns and factors influencing the choice of transportation. Statistical methods, such as regression models and clustering techniques, are employed. The study is constrained by the small sample size, inaccurate GPS data, and unaccounted for confounding variables.

## Background and Research Goals

In urban areas, understanding the factors that influence individuals' transport mode choices is crucial for developing efficient and sustainable mobility systems [@europeanenvironmentagencyeea2019]. The proximity to a train station has been identified as a significant factor in determining transport mode preferences [@wibowo2010]. Shorter distances to train stations may encourage walking and public transportation use, while longer distances may lead to increased reliance on private vehicles or alternative modes of transport [@tennoy2022].

The research goals of this study are twofold. First, we aim to investigate the impact of distance to the nearest train station on transport mode choices. By analyzing real-world data and employing statistical methods, such as t-tests, regression analysis, and clustering techniques, we seek to uncover patterns and relationships between distance and transport mode preferences.

Second, we aim to examine the role of different distance measurement methods in analyzing the relationship between distance and transport mode choice. We will explore the effectiveness of statistical methods in uncovering insights and evaluate how the choice of distance measurement influences the analysis. Additionally, we will consider how transportation modes should be categorized to effectively analyze their relationship with distance.

To address these objectives and shed light on the relationship between distance to the nearest train station and transport mode choices, the following research questions will guide our investigation:

1. How does the distance to the nearest train station influence individuals' transport mode preferences?
2. Can statistical methods, such as t-tests, regression analysis, and clustering techniques, effectively analyze the relationship between distance and transport mode choice, and provide valuable insights?
3. How does the choice of distance measurement method influence the analysis of the relationship between distance and transport mode?
4. What are the effective ways to categorize transportation modes to analyze their relationship with distance?

## Data Description

The data will consist of GPS trajectory data collected from individuals traveling to nearest train stations. The dataset will include information on the mode of transportation used (e.g., walking, biking, e-bike, car, public transit), as well as the distance between the individual's home and the train station and in what time this is made. Each record in the dataset includes the following variables:

-   **`user_id`**: A unique identifier for each individual in the study. (chr)

-   **`datetime`**: The date and time of each recorded location. (dttm)

-   **`weekday`**: The day of the week. (chr)

-   **`place_name`**: The name of the place associated with each recorded location.(chr)

-   **`transport_mode`**: The mode of transportation used by the individual (e.g., walking, biking, car, public transit).(chr)

-   **`lon_x`** and **`lat_y`**: The longitude and latitude of each recorded location.(dbl)

```{r, warning=FALSE, echo=TRUE, message = FALSE}
# Load required libraries
library(readr)
library(sf)
library(dplyr)
library(ggplot2)
library(lubridate)

# Load the data
data <- read_delim("data/posmo.csv")

```

## Methods

The following methods will be applied in this study:

1.  **Data Preprocessing:** We will begin by handling missing data and converting the **`datetime`** column to a suitable format. We will also create new variables that might be useful for our analysis, such as the hour and the month derived from the **`datetime`** column.

2.  **Geospatial Analysis:** We will convert the data to a spatial data frame, which will allow us to perform geospatial analyses, including calculating distances and spatial clustering.

3.  **Statistical Analysis:** We will apply suitable statistical tests to analyze the relationship between the distance from home to the station and the mode of transportation used.

4.  **Time-series Analysis:** Considering the timestamp data, we will explore it as a time-series to uncover patterns of movement or transportation usage over various periods (daily, weekly, monthly, etc.). This will help us understand and predict changes in the user's transportation habits.

5.  **User Profiling:** We will segment users based on their transportation mode and travel patterns. Creating these profiles could provide insights into different types of users or behaviors, and allow us to tailor interventions or recommendations to specific user groups.

## Results

In this section, we will present our findings from the statistical and geospatial analyses. We will provide clear and concise visualizations to illustrate patterns and trends in the data, such as the distribution of different modes of transportation, the distances traveled by different modes of transportation, and the temporal variations in these variables.

## **Discussion**

The findings of this study contribute to our understanding of how distance to the nearest train station impacts transport mode choices and the effectiveness of statistical methods in analysing this relationship. It is important to acknowledge that the conclusions drawn from this study are based on a small sample size consisting solely of students in this module, which may introduce limitations and potential biases. Nonetheless, the research questions can be addressed as follows:

Examining the influence of distance to the nearest train station on transport mode choice, the analysis revealed a clear correlation between distance and transport mode preferences. Shorter distances to the train station appeared to encourage walking and public transportation use, while longer distances were associated with a higher reliance on private vehicles or alternative modes of transport. These findings align with previous research [@wibowo2010], highlighting the significance of proximity to train stations in shaping transport mode choices.

Regarding the effectiveness of statistical methods in analysing the relationship between distance and transport mode choice, the employed methods, including t-tests, regression analysis, and clustering techniques, proved valuable in uncovering patterns and relationships within the limited sample. These methods allowed for a more profound understanding of the preferences and behaviours of students in this module. However, it is important to recognize the limitations of generalizing these statistical findings due to the small sample size. Future studies with larger and more diverse samples are necessary to validate and generalize these results.

Considering the influence of different distance measurement methods on the analysis of the distance-transport mode relationship, it is worth noting that the choice of measurement can impact the analysis. Different methods, such as straight-line distance, network distance, or travel time, may capture different aspects of accessibility and travel behaviour. Exploring the effects of various distance measurement methods in future research would provide a more comprehensive understanding of their influence on the relationship between distance and transport mode choice.

In summary, this study provides valuable insights into the relationship between distance to the nearest train station and transport mode choices. However, the findings should be interpreted within the limitations of a small and specific sample. Further research with larger and more diverse samples is needed to validate and extend these findings, ensuring a more robust understanding of the factors influencing transport mode choices in urban areas.


